---
title: "Data Pre-processing"
subtitle: "Variable selection and normalization"
toc: true
---

The data were pre-processed to prepare them for the subsequent analyses. First, some participants were excluded from the study due to not completing the task or showing very low performance.

After that, a selection of problematic variables was implemented by identifying those with a high amount of missing values and/or high intra-task correlation coefficients. Variables with more than 10% of missing observations were completely removed. After that, correlation analyses were conducted between all cognitive variables inside each task, and statistically significant associations (p < .05) and Person’s r higher than .7 were identified. 

Additionally, a final removal of missing data was performed by excluding all cases with missing values in any of the selected variables. 

Finally, all variables were normalized to z values. Given that two ANT variables (omitted congruent trials and omitted incongruent trials) were inversely correlated with accuracy, they were inverted to facilitate the interpretation of results.


```{r, message= F}

rm(list=ls())
library(tidyverse)
source("helper/helper_tables.R")


base_general <- read.csv("../data/PICT_2014_PRE_filtered.csv")

```


## 1. Variable exclusion

In the first place, we will list all the available variables for each task.

```{r}

variables = colnames(base_general)[-c(1:2)]


# Extraer nombre de la tarea
task <- sub("_.*", "", variables)
task[!task %in% c("ANT", "CORSI", "TOL", "KBIT")] <- "Other"

# Remover el prefijo de tarea (ej: "ANT_") del nombre de variable
clean_name <- sub("^(ANT|CORSI|TOL|KBIT)_", "", variables)
clean_name <- gsub("_", " ", clean_name)

# Crear tabla final
table_vars <- data.frame(Task = task, Variable = clean_name) %>% 
  group_by(Task) %>% 
 summarise(n = n(),
           Variables = paste(Variable, collapse = ", "))

table_vars %>% 
  custom_flextable("Initial variables")






```


### 1.1. Checking variables with missing values

Variables such as RT in incorrect TOL and ANT trials seem to show a high amount of missing values

```{r}

base_general %>% 
  mutate(across(everything(), is.na)) %>% 
  
  mutate(across(everything(), as.numeric)) %>% 
  summarise(across(everything(), sum)) %>% 
  pivot_longer(everything(), names_to = "variables",
               values_to = "perdidos") %>% 
  arrange(desc(perdidos))

```


## 1.2. Checking variables with high correlation values

Correlation analyses between all included variables inside each task were run. In bold, are shown all statistically significant (p \< 0.05) correlations with r \> 0.7.

::: scrolling

```{r}

correlaciones = base_general %>% 
  select(-username,-Year) %>% 
  psych::corr.test(.)


p.correlaciones = correlaciones$p %>% 
  as.data.frame() %>% 
  rownames_to_column("variable_1") %>% 
pivot_longer(cols = -variable_1,
             names_to = "variable_2",
             values_to = "p.value") %>% 
  mutate(p.value = round(p.value, digits = 5))


r.correlaciones = correlaciones$r %>% 
    as.data.frame() %>% 
  rownames_to_column("variable_1") %>% 
pivot_longer(cols = -variable_1,
             names_to = "variable_2",
             values_to = "r")

correlaciones.df = left_join(r.correlaciones, p.correlaciones, by = c("variable_1", "variable_2"))


Correlaciones_tareas <- function(tarea, valor_r, valor_p){
  
  correlaciones.df %>% 
  filter(duplicated(r)) %>% 
  filter(r != 1) %>% 
  filter(r > valor_r) %>%
  filter(p.value < valor_p) %>%
  filter(grepl(tarea, variable_1)) %>% 
  filter(grepl(tarea, variable_2))
  
}

Correlaciones_tareas("ANT", -1, 1) %>% 
  rbind(Correlaciones_tareas("CORSI", -1, 1)) %>% 
  rbind(Correlaciones_tareas("TOL", -1, 1)) %>% 
  rbind(Correlaciones_tareas("KBIT", -1, 1)) %>%
  mutate(Task = gsub("_.*", "", variable_1)) %>%  # Extract task name
  select(Task, everything()) %>%
  mutate(variable_1 =str_remove(variable_1, "^[A-Z]+_")) %>% 
  mutate(variable_2 =str_remove(variable_2, "^[A-Z]+_")) %>%

  custom_flextable("Table 1. Descriptive and inferential analyses for clustering result") %>% 
  bold(j = 1:5,~ p.value < .05 & r > .7) %>%
  valign(valign = "top") 

```

::: 

## 1.3. Selecting variables  

With the selection criteria in mind, determined variables were selected for each task: 

**ANT**
Omitted_trials_amount_congruent, omitted_trials_amount_incongruent, correct_trials_amount_congruent, correct_trials_amount_incongruent, median_rt_correct_congruent, median_rt_correct_incongruent, rt_executive, rt_orienting, rt_alerting 

**CORSI**
Correct_trials_amount, rt_median_correct, rt_median_incorrect 

**TOL**
Correct_trials_amount, rt_planning_median_correct 

**KBIT**
Correct_trials_amount 

```{r}

variables = c(
  
  # ANT
  
  "ANT_omitted_trials_amount_congruent",
  "ANT_omitted_trials_amount_incongruent",
  "ANT_correct_trials_amount_congruent",
  "ANT_correct_trials_amount_incongruent",
  "ANT_median_rt_correct_congruent",
  "ANT_median_rt_correct_incongruent",
  "ANT_rt_ejecutiva",
  "ANT_rt_orientacion",
  "ANT_rt_alerta",
  
  # CORSI
  
  "CORSI_correct_trials_amount",
  "CORSI_rt_median_correct",
  "CORSI_rt_median_incorrect",
  
  # TOL
  
  "TOL_correct_trials_amount",
  "TOL_rt_planning_median_correct",
  
  # KBIT
  
  "KBIT_correct_trials_amount"
  
  )

```

## 2. Data normalization

In order to implement a correct clusterization, all cases with missing values were excluded, and variables were converted to z values.
```{r}


datos.reducidos = base_general %>% 
  select(all_of(c("username",variables))) %>% 
  # select(-starts_with(c("STROOP"))) %>%
  na.omit

write.csv(datos.reducidos, "../data/datos_reducidos.csv", row.names = F) # Includes non scaled data

datos_normalizados <- datos.reducidos %>% 
  select(-username) %>%
  scale() %>% 
  as.data.frame() %>% 
  mutate(across(where(is.numeric), scale)) %>%
  mutate(across((contains("omitted")), ~ . * -1)) ## Variables related to omitted trials were
                                                  ## inverted so positive refer to "better" performaces.
  

write.csv(datos_normalizados, "../data/datos_normalizados.csv", row.names = F) # Includes scaled data

```
